\textls[-0]{\thetitle}

### 2023-03-05 metatheory-condensed.tex
Fix Wsup - Wsup needs specify the index of the constructor.

W-types~\cite{martin1984intuitionistic} are a succinct way to model
inductive types in MLTT.
Together with the identity type $\TyId{\cDot}{\cDot}$, they can express a
whole host of inductive types~\cite{hugunin2021not}.
%
Our enhanced W-type is decided by a W-type signature $\goodWSig{\Gm}{\wsig}{n}$ 
composed of a list (\ruleref{wsig/empty} and \ruleref{wsig/add}) of $n$ pairs of 
types $\{(\wsigproj{i}{1}{\wsig}, \wsigproj{i}{2}{\wsig})\}_i$,
modeling an inductive type of $n$ constructors. 
%
Rule~\ruleref{tm/wsup} states that the $i$-th pair $\goodType{\Gm}{\wsigproj{i}{1}{\wsig}}{}$ and 
$\goodType{\Gm,\wsigproj{i}{1}{\wsig}}{\wsigproj{i}{2}{\wsig}}{}$ models the $i$-th constructor---given 
two arguments $\goodTerm{\Gm}{t_1}{\wsigproj{i}{1}{\wsig}}$ and
$\goodTerm{\Gm,\sub{\wsigproj{i}{2}{\wsig}}{\SubstExt{\SubstWeak{0}}{t_1}}}{t_2}{\El{\wcode{\wsig}}}$,
we can construct a term $\Wsup{\wsig}{t_1}{t_2}$ of the W-type. $\wcode{\wsig}$ gives the code of the
W-type (\ruleref{tm/w}).

@@

\begin{figure}
  \begin{tabular}{
    >{\columncolor[HTML]{C0C0C0}}l lcl}
    \hline
    \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}surface syntax} &
      \multicolumn{1}{l|}{Equivalent Type} &
      \multicolumn{1}{l}{} &
      \multicolumn{1}{l|}{} \\
    \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}\lsti!| unit : tm!} &
      \multicolumn{1}{l|}{$\bbone \times (\bbzero \to tm) \to tm$} &
      $\tau'$ &
      \multicolumn{1}{l|}{$\coloneqq\WSigAdd{\WSigEmp}{\bbone}{\bbzero}$} \\
    \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}\lsti!| app  : tm -> tm -> tm!} &
      \multicolumn{1}{l|}{$\bbone \times (\bbtwo \to tm) \to tm$} &
      $\tau$ &
      \multicolumn{1}{l|}{\cellcolor[HTML]{FFFFC7}$\coloneqq\WSigAdd{\tau'}{\bbone}{\bbtwo}$} \\ \hline
    \multicolumn{1}{c}{\cellcolor[HTML]{C0C0C0}\lsti!unit!} &
      $\coloneqq \Wsup{\tau}{\unit}{"elim-"\bbzero}$ &
      \cellcolor[HTML]{C0C0C0}\lsti!app unit unit! &
      $\coloneqq\Wsup{\tau}{\unit}{"unit"}$
    \end{tabular}
\end{figure}

@@

In contrast, $\wsig$ should be hidden from the typing context of any
term that does not invoke $\Wsup{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$, so that
the term can be reused---without being rechecked---for a different
W-type signature~$\wsig'$ that extends~$\wsig$ with additional
constructors.
Moreover, the typing of the term should be made parametric to
the definitions of those fields that invoke $\texttt{Wsup}$ or $\texttt{Wrec}$,
so that the term can be reused---without being rechecked---when those fields
are overridden to support the extended signature~$\wsig'$.
%
Such abstraction required by family polymorphism is supported via linkages,
which we discuss next.

@@

%recorded by the first premise $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$
The type~$A$ in \ruleref{l/add} and \ruleref{lsig/add},
abstracting the types of the prior fields, does not necessarily contain the same types as those recorded by~$\TyLkg{\lsig}$,
%\EDJ{Suggest changing to ``not necessarily the same as those recorded by lsig in the prior fields''},
because a field defined as the code $\wcode{\wsig}$ of a W-type has to
expose different types to different fields that come after it.
Later fields that invoke
$\Wsup{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$
should see the concrete signature~$\wsig$, as rules \ruleref{tm/wsup}
and \ruleref{tm/wrec} stipulate.
By contrast, $\wsig$ should be hidden
from all other fields, so that they can be reused
in a different context where the W-type signature~$\wsig$ is replaced by an extended one~$\wsig'$.

@@

\item 
The next four fields model the four constructors of~\lsti{tm}.
Constructor \lsti{tm_unit} is modeled as $\Wsup{\wsig_{\texttt{tm}}}{\top}{\bot}$
and typed with $\El{\mathit{self}.\texttt{tm}}$, where $\mathit{self}$ stands for
the typing context containing the first field~\lsti{tm}.
The W-type signature $\wsig_{\texttt{tm}}$ is exposed in
this typing context; that is, \lsti{tm} has type $\TyS{\wcode{\wsig_{\texttt{tm}}}}$.
So $\El{\mathit{self}.\texttt{tm}}$ and $\El{\wcode{\wsig_{\texttt{tm}}}}$ can be equated,
as required by rule \ruleref{tm/wsup}.

@@

{\goodTerm{\Gm}{\Wsup{T}{a}{b}}{\El{(\wcode{\wsig})}} }

@@

{\goodTerm{\Gamma}{\Wrec{\wsig}{h}{(\Wsup{\wsig}{a}{b})} \equiv 

>>>>

W-types~\cite{martin1984intuitionistic} are a succinct way to model
inductive types in MLTT.
Together with the identity type $\TyId{\cDot}{\cDot}$, they can express a
whole host of inductive types~\cite{hugunin2021not}.
%
Our enhanced W-type is decided by a W-type signature $\goodWSig{\Gm}{\wsig}{n}$ 
composed of a list (\ruleref{wsig/empty} and \ruleref{wsig/add}) of $n$ pairs of 
types $\{(\wsigproj{i}{1}{\wsig}, \wsigproj{i}{2}{\wsig})\}_i$,
modeling an inductive type of $n$ constructors. 
%
Rule~\ruleref{tm/wsup} states that the $i$-th pair $\goodType{\Gm}{\wsigproj{i}{1}{\wsig}}{}$ and 
$\goodType{\Gm,\wsigproj{i}{1}{\wsig}}{\wsigproj{i}{2}{\wsig}}{}$ models the $i$-th constructor---given 
two arguments $\goodTerm{\Gm}{t_1}{\wsigproj{i}{1}{\wsig}}$ and
$\goodTerm{\Gm,\sub{\wsigproj{i}{2}{\wsig}}{\SubstExt{\SubstWeak{0}}{t_1}}}{t_2}{\El{\wcode{\wsig}}}$,
we can construct a term $\Wsup{i}{\wsig}{t_1}{t_2}$ of the W-type. $\wcode{\wsig}$ gives the code of the
W-type (\ruleref{tm/w}).

@@

\begin{figure}
  \begin{tabular}{
    >{\columncolor[HTML]{C0C0C0}}l lcl}
    \hline
    \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}surface syntax} &
      \multicolumn{1}{l|}{Equivalent Type} &
      \multicolumn{1}{l}{} &
      \multicolumn{1}{l|}{} \\
    \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}\lsti!| unit : tm!} &
      \multicolumn{1}{l|}{$\bbone \times (\bbzero \to tm) \to tm$} &
      $\tau'$ &
      \multicolumn{1}{l|}{$\coloneqq\WSigAdd{\WSigEmp}{\bbone}{\bbzero}$} \\
    \multicolumn{1}{|l}{\cellcolor[HTML]{C0C0C0}\lsti!| app  : tm -> tm -> tm!} &
      \multicolumn{1}{l|}{$\bbone \times (\bbtwo \to tm) \to tm$} &
      $\tau$ &
      \multicolumn{1}{l|}{\cellcolor[HTML]{FFFFC7}$\coloneqq\WSigAdd{\tau'}{\bbone}{\bbtwo}$} \\ \hline
    \multicolumn{1}{c}{\cellcolor[HTML]{C0C0C0}\lsti!unit!} &
      $\coloneqq \Wsup{1}{\tau}{\unit}{"elim-"\bbzero}$ &
      \cellcolor[HTML]{C0C0C0}\lsti!app unit unit! &
      $\coloneqq\Wsup{0}{\tau}{\unit}{"unit"}$
    \end{tabular}
\end{figure}

@@

In contrast, $\wsig$ should be hidden from the typing context of any
term that does not invoke $\Wsup{}{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$, so that
the term can be reused---without being rechecked---for a different
W-type signature~$\wsig'$ that extends~$\wsig$ with additional
constructors.
Moreover, the typing of the term should be made parametric to
the definitions of those fields that invoke $\texttt{Wsup}$ or $\texttt{Wrec}$,
so that the term can be reused---without being rechecked---when those fields
are overridden to support the extended signature~$\wsig'$.
%
Such abstraction required by family polymorphism is supported via linkages,
which we discuss next.

@@

%recorded by the first premise $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$
The type~$A$ in \ruleref{l/add} and \ruleref{lsig/add},
abstracting the types of the prior fields, does not necessarily contain the same types as those recorded by~$\TyLkg{\lsig}$,
%\EDJ{Suggest changing to ``not necessarily the same as those recorded by lsig in the prior fields''},
because a field defined as the code $\wcode{\wsig}$ of a W-type has to
expose different types to different fields that come after it.
Later fields that invoke
$\Wsup{}{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$
should see the concrete signature~$\wsig$, as rules \ruleref{tm/wsup}
and \ruleref{tm/wrec} stipulate.
By contrast, $\wsig$ should be hidden
from all other fields, so that they can be reused
in a different context where the W-type signature~$\wsig$ is replaced by an extended one~$\wsig'$.

@@

\item 
The next four fields model the four constructors of~\lsti{tm}.
Constructor \lsti{tm_unit} is modeled as $\Wsup{}{\wsig_{\texttt{tm}}}{\top}{\bot}$
and typed with $\El{\mathit{self}.\texttt{tm}}$, where $\mathit{self}$ stands for
the typing context containing the first field~\lsti{tm}.
The W-type signature $\wsig_{\texttt{tm}}$ is exposed in
this typing context; that is, \lsti{tm} has type $\TyS{\wcode{\wsig_{\texttt{tm}}}}$.
So $\El{\mathit{self}.\texttt{tm}}$ and $\El{\wcode{\wsig_{\texttt{tm}}}}$ can be equated,
as required by rule \ruleref{tm/wsup}.

@@

{\goodTerm{\Gm}{\Wsup{}{T}{a}{b}}{\El{(\wcode{\wsig})}} }

@@

{\goodTerm{\Gamma}{\Wrec{\wsig}{h}{(\Wsup{}{\wsig}{a}{b})} \equiv 
